---
title: "Group 6 Kaggle Project"
author: "Group 6"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(mgcv)
library(rpart)
library(mlr3)
```

## Load the data

```{r}
survey <- read_csv("survey_train_test.csv")
survey |> count(test)
```

There are 3,027 training survey rows and 2,000 test survey rows.

```{r}
col_select = c(
    "Order Date",
    "Purchase Price Per Unit",
    "Quantity",
    "Shipping Address State",
    "Category",
    "Survey ResponseID"
  )

purchases <- read_csv(
  "amazon-purchases.csv",
  col_select = all_of(col_select),
  col_types = cols(
    `Order Date` = col_date(format = "%Y-%m-%d"),
    `Purchase Price Per Unit` = col_double(),
    Quantity = col_double(),
    `Shipping Address State` = col_character(),
    Category = col_character(),
    `Survey ResponseID` = col_character()
  )
) |>
  rename(
    date = `Order Date`,
    price = `Purchase Price Per Unit`,
    quantity = Quantity,
    ship_state = `Shipping Address State`,
    category = Category,
    id = `Survey ResponseID`
  )
```

## Explore the data

```{r}
survey |>
  filter(!test) |>
  ggplot(aes(x = factor(Q.amazon.use.hh.size.num))) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Household size",
    y = "Number of people",
    title = "Training response"
  ) 
```

Household size is coded as 1, 2, 3, or 4, where 4 means four or more people.

## Clean the data

The dataset paper studies purchases from January 2018 through October 2022 because data collection began in November 2022. Therefore, October 2022 is the last complete month covered for all participants. We will use the same end date.

```{r}
purchases <- purchases |>
  mutate(
    category = replace_na(category, "MISSING")
  ) |>
  filter(date <= as.Date("2022-10-31")) |>
  mutate(
    spend = price * quantity,
    weekend = format(date, "%u") %in% c("6", "7"),
    baby = str_detect(
      category,
      "BABY|INFANT|TODDLER|PACIFIER|TEETHER|BREAST_PUMP|BOTTLE_NIPPLE|STROLLER|CRIB"
    ),
    toy = str_detect(category, "TOY|PUZZLE|DOLL") & category != "PET_TOY"
  )

features <- purchases |>
  group_by(id) |>
  summarize(
    n_purchase = n(),
    total_spend = sum(spend, na.rm = TRUE),
    n_category = n_distinct(category[category != "MISSING"]),
    n_state = n_distinct(ship_state, na.rm = TRUE),
    active_days = as.numeric(max(date) - min(date)) + 1,
    prop_weekend = mean(weekend),
    prop_missing_category = mean(category == "MISSING"),
    n_baby = sum(baby),
    n_toy = sum(toy),
    .groups = "drop"
  ) |>
  mutate(
    log_n_purchase = log1p(n_purchase),
    log_total_spend = log1p(total_spend),
    log_n_category = log1p(n_category),
    log_active_days = log1p(active_days),
    log_n_baby = log1p(n_baby),
    log_n_toy = log1p(n_toy)
  ) |>
  select(
    id, log_n_purchase, log_total_spend, log_n_category,
    log_active_days, n_state, prop_weekend, prop_missing_category,
    log_n_baby, log_n_toy
  )
```

## Modeling

Only gender and state are observed for both the training and test survey rows.
States with fewer than 30 training rows are combined into an `Other` group.

```{r}
survey_small <- survey |>
  mutate(
    id = Survey.ResponseID,
    y = Q.amazon.use.hh.size.num,
    gender = Q.demos.gender,
    state = Q.demos.state
  ) |>
  select(id, y, gender, state, test)

keep_state <- survey_small |>
  filter(!test) |>
  count(state) |>
  filter(n >= 30) |>
  pull(state)

survey_small <- survey_small |>
  mutate(
    gender = ifelse(
      gender %in% c("Female", "Male"),
      as.character(gender),
      "Other"
    ),
    state = ifelse(
      state %in% keep_state,
      as.character(state),
      "Other"
    ),
    gender = factor(gender),
    state = factor(state)
  )

feature_names <- setdiff(names(features), "id")

model_data <- survey_small |>
  left_join(features, by = "id") |>
  mutate(across(all_of(feature_names), ~ replace_na(.x, 0)))

train_data <- model_data |>
  filter(!test) |>
  select(-id, -test)

test_data <- model_data |>
  filter(test)
```

## Predictions

```{r}
lm_formula <- y ~ gender + state + log_n_purchase + log_total_spend +
  log_n_category + log_active_days + n_state + prop_weekend +
  prop_missing_category + log_n_baby + log_n_toy

gam_formula <- y ~ gender + state +
  s(log_n_purchase, k = 5) +
  s(log_total_spend, k = 5) +
  s(log_n_category, k = 5) +
  s(log_active_days, k = 5) +
  s(n_state, k = 4) +
  s(log_n_baby, k = 5) +
  s(log_n_toy, k = 5) +
  prop_weekend +
  prop_missing_category

rmse <- function(truth, prediction) {
  sqrt(mean((truth - prediction)^2))
}
```

```{r}
set.seed(101)

task <- as_task_regr(train_data, target = "y")

cv_10fold <- rsmp("cv", folds = 10)
cv_10fold$instantiate(task)

k <- cv_10fold$iters

model_names <- c(
  "Baseline", "Linear regression", "GAM", "Regression tree"
)
cv_error <- matrix(NA, nrow = k, ncol = length(model_names))

for (i in 1:k) {
  train_fold <- train_data[cv_10fold$train_set(i), ]
  valid_fold <- train_data[cv_10fold$test_set(i), ]

  mean_prediction <- rep(mean(train_fold$y), nrow(valid_fold))

  lm_fit <- lm(lm_formula, data = train_fold)
  lm_prediction <- predict(lm_fit, newdata = valid_fold)

  gam_fit <- mgcv::gam(gam_formula, data = train_fold, method = "REML")
  gam_prediction <- predict(gam_fit, newdata = valid_fold)

  tree_fit <- rpart(lm_formula, data = train_fold, method = "anova")
  tree_prediction <- predict(tree_fit, newdata = valid_fold)

  predictions <- list(
    mean_prediction,
    lm_prediction,
    gam_prediction,
    tree_prediction
  )

  for (j in 1:length(model_names)) {
    cv_error[i, j] <- rmse(valid_fold$y, predictions[[j]])
  }
}

cv_results <- tibble(
  model = factor(model_names, levels = model_names),
  rmse = colMeans(cv_error)
)

cv_results
```

```{r}
cv_results |>
  ggplot(aes(x = model, y = rmse)) +
  geom_col(fill = "steelblue") +
  coord_cartesian(ylim = c(1, max(cv_results$rmse) + 0.01)) +
  labs(
    x = NULL,
    y = "10-fold CV RMSE",
    title = "Model comparison"
  )
```

The GAM has the smallest cross-validation RMSE.

Let's try some additional models!

```{r}
library("mlr3verse")

mod_mat <- model.matrix(~ ., data = train_data)
num_data <- as.data.frame(mod_mat[, -1])

ridge_fit <- lrn(
  "regr.cv_glmnet",
  alpha = 0,
  nfolds = 10,
  s = "lambda.min",
  lambda = 10^seq(from = -1.5, to = 1.5, by = 0.1),
  standardize = TRUE
)

lasso_fit <- lrn(
  "regr.cv_glmnet",
  alpha = 1,
  nfolds = 10,
  s = "lambda.min",
  lambda = 10^seq(from = -2, to = 2, by = 0.1),
  standardize = TRUE
)

rf_fit <- lrn("regr.ranger", importance = "impurity")

xgb_fit <- as_learner(
  po("encode", method = "one-hot") %>>% lrn("regr.xgboost")
)
```

```{r}
set.seed(101)

new_cv_error <- matrix(NA, nrow = k, ncol = 4)

for (i in 1:k) {
  train_id <- cv_10fold$train_set(i)
  valid_id <- cv_10fold$test_set(i)

  train_fold <- train_data[train_id, ]
  valid_fold <- train_data[valid_id, ]
  train_num <- num_data[train_id, ]
  valid_num <- num_data[valid_id, ]

  train_task <- as_task_regr(train_fold, target = "y")
  num_task <- as_task_regr(train_num, target = "y")

  ridge_fit$train(task = num_task)
  ridge_pred <- ridge_fit$predict_newdata(newdata = valid_num)$response

  lasso_fit$train(task = num_task)
  lasso_pred <- lasso_fit$predict_newdata(newdata = valid_num)$response

  rf_fit$train(task = train_task)
  rf_pred <- rf_fit$predict_newdata(newdata = valid_fold)$response

  xgb_fit$train(task = train_task)
  xgb_pred <- xgb_fit$predict_newdata(newdata = valid_fold)$response

  new_cv_error[i, ] <- c(
    rmse(valid_fold$y, ridge_pred),
    rmse(valid_fold$y, lasso_pred),
    rmse(valid_fold$y, rf_pred),
    rmse(valid_fold$y, xgb_pred)
  )
}
```

```{r}
new_model_names <- c(
  "GAM", "Ridge", "Lasso", "Random Forest", "XGBoost"
)

gam_rmse <- cv_results$rmse[cv_results$model == "GAM"]
new_rmse <- colMeans(new_cv_error)

model_results <- data.frame(
  model = factor(new_model_names, levels = new_model_names),
  rmse = round(c(gam_rmse, new_rmse), 6)
)

model_results
```

```{r}
model_results |>
  ggplot(aes(x = model, y = rmse)) +
  geom_col(fill = "steelblue") +
  coord_cartesian(ylim = c(1, max(model_results$rmse) + 0.01)) +
  labs(
    x = NULL,
    y = "10-fold CV RMSE",
    title = "Model comparison"
  )
```

The GAM is still the winner!

Let's try adding some new features.

```{r}
product_features_v2 <- read_csv(
  "amazon-purchases.csv",
  col_select = c(
    "Order Date", "ASIN/ISBN (Product Code)", "Survey ResponseID"
  ),
  col_types = cols(
    `Order Date` = col_date(format = "%Y-%m-%d"),
    `ASIN/ISBN (Product Code)` = col_character(),
    `Survey ResponseID` = col_character()
  )
) |>
  rename(
    date = `Order Date`,
    product = `ASIN/ISBN (Product Code)`,
    id = `Survey ResponseID`
  ) |>
  filter(
    date <= as.Date("2022-10-31"),
    !is.na(product)
  ) |>
  group_by(id) |>
  summarize(
    n_product = n_distinct(product),
    n_product_rows = n(),
    .groups = "drop"
  ) |>
  mutate(
    log_n_product = log1p(n_product),
    prop_repeat = 1 - n_product / n_product_rows
  ) |>
  select(id, log_n_product, prop_repeat)

category_features_v2 <- purchases |>
  filter(category != "MISSING") |>
  count(id, category, name = "category_n") |>
  group_by(id) |>
  summarize(
    top_category_prop = max(category_n) / sum(category_n),
    .groups = "drop"
  )

features_v2 <- purchases |>
  mutate(
    month = format(date, "%Y-%m"),
    food = category %in% c("FOOD", "GROCERY", "VEGETABLE", "FRUIT"),
    health = category %in% c(
      "HEALTH_PERSONAL_CARE", "MEDICATION", "VITAMIN"
    ),
    pet = category %in% c("PET_FOOD", "PET_SUPPLIES")
  ) |>
  group_by(id) |>
  summarize(
    active_months = n_distinct(month),
    mean_price = mean(price, na.rm = TRUE),
    mean_quantity = mean(quantity, na.rm = TRUE),
    prop_multi = mean(quantity > 1, na.rm = TRUE),
    purchases_per_month = n() / active_months,
    purchase_days = n_distinct(date),
    prop_baby = mean(baby),
    prop_toy = mean(toy),
    n_food = sum(food),
    prop_food = mean(food),
    n_health = sum(health),
    prop_health = mean(health),
    n_pet = sum(pet),
    prop_pet = mean(pet),
    .groups = "drop"
  ) |>
  mutate(
    log_active_months = log1p(active_months),
    log_mean_price = log1p(mean_price),
    log_purchases_per_month = log1p(purchases_per_month),
    log_purchase_days = log1p(purchase_days),
    log_n_food = log1p(n_food),
    log_n_health = log1p(n_health),
    log_n_pet = log1p(n_pet)
  ) |>
  select(
    id, log_active_months, log_mean_price, mean_quantity,
    prop_multi, log_purchases_per_month, log_purchase_days,
    prop_baby, prop_toy, log_n_food, prop_food,
    log_n_health, prop_health, log_n_pet, prop_pet
  ) |>
  left_join(product_features_v2, by = "id") |>
  left_join(category_features_v2, by = "id")

new_names <- setdiff(names(features_v2), "id")

model_data_v2 <- model_data |>
  left_join(features_v2, by = "id") |>
  mutate(across(all_of(new_names), ~ replace_na(.x, 0)))

train_data_v2 <- model_data_v2 |>
  filter(!test) |>
  select(-id, -test)
```

```{r}
gam_formula_v2 <- y ~ gender + state +
  s(log_n_purchase, k = 5) +
  s(log_total_spend, k = 5) +
  s(log_n_category, k = 5) +
  s(log_active_days, k = 5) +
  s(n_state, k = 4) +
  s(log_n_baby, k = 5) +
  s(log_n_toy, k = 5) +
  prop_weekend + prop_missing_category +
  log_active_months + log_mean_price + mean_quantity +
  prop_multi + log_purchases_per_month + log_purchase_days +
  log_n_product + prop_repeat + top_category_prop +
  prop_baby + prop_toy + log_n_food + prop_food +
  log_n_health + prop_health + log_n_pet + prop_pet

mod_mat_v2 <- model.matrix(~ ., data = train_data_v2)
num_data_v2 <- as.data.frame(mod_mat_v2[, -1])

ridge_fit_v2 <- lrn(
  "regr.cv_glmnet",
  alpha = 0,
  nfolds = 10,
  s = "lambda.min",
  lambda = 10^seq(from = -1.5, to = 1.5, by = 0.1),
  standardize = TRUE
)

lasso_fit_v2 <- lrn(
  "regr.cv_glmnet",
  alpha = 1,
  nfolds = 10,
  s = "lambda.min",
  lambda = 10^seq(from = -2, to = 2, by = 0.1),
  standardize = TRUE
)

rf_fit_v2 <- lrn("regr.ranger", importance = "impurity")

xgb_fit_v2 <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn(
      "regr.xgboost",
      max_depth = 2,
      eta = 0.01,
      nrounds = 1000,
      min_child_weight = 30,
      subsample = 0.75,
      colsample_bytree = 0.65,
      lambda = 15,
      alpha = 0.5
    )
)
```

```{r}
set.seed(101)

v2_cv_error <- matrix(NA, nrow = k, ncol = 5)

for (i in 1:k) {
  train_id <- cv_10fold$train_set(i)
  valid_id <- cv_10fold$test_set(i)

  train_fold_v2 <- train_data_v2[train_id, ]
  valid_fold_v2 <- train_data_v2[valid_id, ]
  train_num_v2 <- num_data_v2[train_id, ]
  valid_num_v2 <- num_data_v2[valid_id, ]

  train_task_v2 <- as_task_regr(train_fold_v2, target = "y")
  num_task_v2 <- as_task_regr(train_num_v2, target = "y")

  gam_fit_v2 <- mgcv::gam(
    gam_formula_v2,
    data = train_fold_v2,
    method = "REML"
  )
  gam_pred_v2 <- predict(gam_fit_v2, newdata = valid_fold_v2)

  ridge_fit_v2$train(task = num_task_v2)
  ridge_pred_v2 <- ridge_fit_v2$predict_newdata(
    newdata = valid_num_v2
  )$response

  lasso_fit_v2$train(task = num_task_v2)
  lasso_pred_v2 <- lasso_fit_v2$predict_newdata(
    newdata = valid_num_v2
  )$response

  rf_fit_v2$train(task = train_task_v2)
  rf_pred_v2 <- rf_fit_v2$predict_newdata(
    newdata = valid_fold_v2
  )$response

  xgb_fit_v2$train(task = train_task_v2)
  xgb_pred_v2 <- xgb_fit_v2$predict_newdata(
    newdata = valid_fold_v2
  )$response

  v2_cv_error[i, ] <- c(
    rmse(valid_fold_v2$y, gam_pred_v2),
    rmse(valid_fold_v2$y, ridge_pred_v2),
    rmse(valid_fold_v2$y, lasso_pred_v2),
    rmse(valid_fold_v2$y, rf_pred_v2),
    rmse(valid_fold_v2$y, xgb_pred_v2)
  )
}
```

```{r}
v2_names <- c(
  "GAM", "GAM v2", "Ridge v2", "Lasso v2",
  "Random Forest v2", "XGBoost v2"
)

v2_rmse <- colMeans(v2_cv_error)

model_results_v2 <- data.frame(
  model = factor(v2_names, levels = v2_names),
  rmse = round(c(gam_rmse, v2_rmse), 6)
)

model_results_v2
```

```{r}
model_results_v2 |>
  ggplot(aes(x = model, y = rmse)) +
  geom_col(fill = "steelblue") +
  coord_cartesian(ylim = c(1, max(model_results_v2$rmse) + 0.01)) +
  labs(
    x = NULL,
    y = "10-fold CV RMSE",
    title = "Model comparison"
  )
```

XGBoost v2 has the lowest CV RMSE.

## Fit the final model and make the submission

```{r}
set.seed(101)

test_data_v2 <- model_data_v2 |>
  filter(test) |>
  select(-id, -test, -y)

final_task <- as_task_regr(train_data_v2, target = "y")

xgb_fit_v2$train(task = final_task)

test_prediction <- xgb_fit_v2$predict_newdata(
  newdata = test_data_v2
)$response

submission <- read_csv(
  "sample_sol.csv"
)

stopifnot(identical(submission$Survey.ResponseID, test_data$id))

submission$Q.amazon.use.hh.size.num <- test_prediction

write_csv(submission, "group_6_submission.csv")

head(submission)
```
